---
title: "Analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

This is our analysis of the [Science Museum of Minnesota](https://www.smm.org) member data for participation in the [Analyze This!](http://www.meetup.com/AnalyzeThis/) 2016 Science Museum Summer Challenge.

### Keep It Statistically Simple Team

- Justin Grammens <justin@recursiveawesome.com>
- Tron Kotz <tron.kotz@gmail.com>
- Lance Kaiwei <lancekaiwei@gmail.com>
- Saad Khan <saadkhan321@gmail.com>
- Peter Edstrom <peter@softwareforgood.com>


## Import the data into dataframes

Let's start by pulling the data into local variables. Noted here so you can see where the data comes from. If you do not have a copy of the `SMM Competitor Workbook 5.25.16.xlsx` file, please contact Beth Varro at <bvarro@smm.org> and fill out an NDA.

```{r data}
library(readxl)
workbook.file <- "data/SMM Competitor Workbook 5.25.16.xlsx"
ad <- read_excel(workbook.file, sheet="Analytics Dataset")
ml <- read_excel(workbook.file, sheet="Mbrshp Level")
pm <- read_excel(workbook.file, sheet="Pmt Mthd")
sc <- read_excel(workbook.file, sheet="Sales Channel")
o1 <- read_excel(workbook.file, sheet="Offer on 1st Ren Notice")
ei <- read_excel(workbook.file, sheet="Email Indicators")
eo <- read_excel(workbook.file, sheet="EMail Options")
d1 <- read_excel(workbook.file, sheet="Demo_1")
zc <- read_excel(workbook.file, sheet="Zip Code")
kp <- read_excel(workbook.file, sheet="Key people on Membership")
nv <- read_excel(workbook.file, sheet="Num Visits")
cs <- read_excel(workbook.file, sheet="CSIs")
co <- read_excel(workbook.file, sheet="Communication")
```

## Basic Visuals

Let's take a first pass to see what the data looks like. 

```{r ad prep}
library(ggplot2)
ad$`Initiation Month` <- cut(ad$`Initiation Date`, breaks="month")
ad$`Expiration Month` <- cut(ad$`Expiration Date`, breaks="month")
```

Here we have a bar chart showing in what months the initiations are found.

```{r ad chart for Initiations, echo=FALSE, fig.height=2}
ggplot(data=ad,
  aes(`Initiation Month`,1)) +
  stat_summary(fun.y=sum, geom="bar") +
  theme(axis.text.x=element_text(angle = 45, hjust = 1), axis.title.x=element_blank()) +
  ylab("Initiation Count")
```

And this shows the expirations by month. Showing expirations as a negative count.

```{r ad chart for Expiration, echo=FALSE, fig.height=2}
ggplot(data=ad,
  aes(`Expiration Month`,-1)) +
  stat_summary(fun.y=sum, geom="bar") +
  theme(axis.text.x=element_text(angle = 45, hjust = 1), axis.title.x=element_blank()) +
  ylab("Expiration Count")
```


